### 2018-02-23 Merow2014_Appendix.Rmd
Appendix E - good & bad years; Appendix F - environmental covariate

---
title: "Merow 2014 Appendix"
author: "Tim Szewczyk"
date: "2/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkgs <- c("tidyverse", "magrittr", "fields", "IPMpack")
suppressMessages(invisible(lapply(pkgs, library, character.only=TRUE)))
theme_set(theme_bw())
```

```{r App_A}
d <- read.csv("data/Appendix_A_Simple_data.csv")
# Appendix A from Merow 2014
########
## Fit regressions
########
  p <- NULL
  # survival logistic regression
  surv.reg <- glm(surv ~ size, family="binomial", data=d)
  p$surv.int <- coef(surv.reg)[1]
  p$surv.slope <- coef(surv.reg)[2]

  # growth linear regression
  growth.reg <- gls(sizeNext ~ size + I(size^2), weights=varExp(), 
                    na.action=na.omit, data=d)
  p$growth.int <- coef(growth.reg)[1]
  p$growth.slope <- coef(growth.reg)[2]
  p$growth.sqrd <- coef(growth.reg)[3]
  p$growth.sigma2 <- summary(growth.reg)$sigma^2
  p$growth.sigma2.exp <- as.numeric(growth.reg$modelStruct$varStruct)

  # flower logistic regression
  flower.reg <- glm(fec.flower ~ size, family="binomial", data=d)
  p$flower.int <- coef(flower.reg)[1]
  p$flower.slope <- coef(flower.reg)[2]

  # seed poisson regression
  seed.reg <- glm(fec.seed ~ size, family="poisson", data=d[d$fec.flower==1,])
  p$seed.int <- coef(seed.reg)[1]
  p$seed.slope <- coef(seed.reg)[2]

  # recruit size distribution
  p$recruit.size.mean <- mean(d$sizeNext[is.na(d$size)])
  p$recruit.size.sd <- sd(d$sizeNext[is.na(d$size)])

  # establishment probability
  p$establishment.prob <- sum(is.na(d$size))/sum(d$fec.seed, na.rm=TRUE)

  par(mfrow=c(2,2), mar=c(4,4,2,1))
  xx <- data.frame(size=seq(0, 8, by=.01))
  plot(surv ~ size, data=d, col=rgb(0,0,0,0.5), 
       xlab="Size (t)", ylab="Survival to t+1")
  lines(xx$size, predict(surv.reg, xx, type="response"), col="red")
  plot(sizeNext ~ size, data=d, col=rgb(0,0,0,0.5), 
       xlab="Size (t)", ylab="Size (t+1)")
  lines(xx$size, predict(growth.reg, xx), col="red")
  plot(fec.seed ~ size, data=d, col=rgb(0,0,0,0.5), 
       xlab="Size (t)", ylab="Seed number")
  lines(xx$size, predict(seed.reg, xx, type="response"), col="red")
  lines(xx$size, predict(flower.reg, xx, type="response")*
          max(d$fec.seed, na.rm=TRUE), col="blue")
  hist(d$sizeNext[is.na(d$size)], xlab="Recruit size", main="", freq=FALSE)
  lines(xx$size, dnorm(xx$size, p$recruit.size.mean, p$recruit.size.sd),
        col="red")

########
## Life history functions
########
  # survival probability
  s.x <- function(x, p) {
    u <- exp(p$surv.int + p$surv.slope*x)
    return(u / (1+u))
  }

  # growth
  g.yx <- function(xp, x, p) {
    dnorm(xp, p$growth.int + p$growth.slope*x + p$growth.sqrd*x^2, 
          sqrt(p$growth.sigma2 * exp(2*p$growth.sigma2.exp*x)))
  }

  # reproduction
  p.flower.x <- function(x, p) {
    u <- exp(p$flower.int + p$flower.slope*x)
    return(u/(1+u))
  }
  f.yx <- function(xp, x, p) {
    p.flower.x(x, p)*
      p$establishment.prob*
      dnorm(xp, p$recruit.size.mean, p$recruit.size.sd)*
      exp(p$seed.int + p$seed.slope*x)
  }

########
## Kernel creation
########
  min.size <- .9*min(c(d$size, d$sizeNext), na.rm=TRUE)
  max.size <- 1.1*max(c(d$size, d$sizeNext), na.rm=TRUE)
  n <- 100  # ncells in matrix
  b <- min.size + c(0:n)*(max.size - min.size)/n  # boundary points
  y <- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
  h <- y[2] - y[1]  # step size

  G.mx <- h*outer(y, y, g.yx, p=p)
  S.v <- s.x(y, p)
  F.mx <- h*outer(y, y, f.yx, p=p)
  P.mx <- G.mx
  # Fix eviction of individuals below/above size limits
  for(i in 1:(n/2)) {
    G.mx[1,i] <- G.mx[1,i] + 1 - sum(G.mx[,i])
    P.mx[,i] <- G.mx[,i]*S.v[i]
  }
  for(i in (n/2+1):n) {
    G.mx[n,i] <- G.mx[n,i] + 1 - sum(G.mx[,i])
    P.mx[,i] <- G.mx[,i]*S.v[i]
  }
  K.mx <- P.mx + F.mx

########
## Summaries & plots
########
  lam <- Re(eigen(K.mx)$values[1])
  w.eigen <- Re(eigen(K.mx)$vectors[,1])
  stable.dist <- w.eigen/sum(w.eigen)
  v.eigen <- Re(eigen(t(K.mx))$vectors[,1])
  repro.val <- v.eigen/v.eigen[1]
  v.dot.w <- sum(stable.dist*repro.val)*h
  sens <- outer(repro.val, stable.dist)/v.dot.w
  elas <- matrix(as.vector(sens)*as.vector(K.mx)/lam, nrow=n)

  par(mfrow=c(2,3), mar=c(4,5,2,2))
  image.plot(y, y, t(K.mx),  
             xlab="Size (t)", ylab="Size (t+1)", main="IPM matrix")
  abline(a=0, b=1, lty=2)
  contour(y, y, t(K.mx), add=TRUE, drawlabels=TRUE)
  image.plot(y, y, t(elas), 
             xlab="Size (t)", ylab="Size (t+1)", main="Elasticity")
  image.plot(y, y, t(sens), 
             xlab="Size (t)", ylab="Size (t+1)", main="Sensitivity")
  plot(y, stable.dist, xlab="Size", type="l", main="Stable size distribution")
  plot(y, repro.val, xlab="Size", type="l", main="Reproductive values")
  plot(y, s.x(y, p), xlab="Size", ylab="Pr(surv)", main="Eviction")
  lines(y, apply(P.mx, 2, sum), col="red")
```

```{r App_B}
data("dataIPMpackSuccisa")
Sp1 <- dataIPMpackSuccisa

par(mfrow=c(3,2), mar=c(4,4,2,1)) 
plot(Sp1$size, Sp1$surv, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Survival to t+1")
plot(Sp1$size, Sp1$sizeNext, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Size (t+1)") 
plot(Sp1$size, Sp1$fec1Bolt, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Flowering probability")
plot(Sp1$size, Sp1$fec2Stem, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Stem Number")
plot(Sp1$size, Sp1$fec3Head, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Flower Head Number") 
hist(Sp1$sizeNext[is.na(Sp1$size)], main="", xlab="Recruit Size")

# size ranges
x <- seq(0, 10, length.out=1001)
x0 <- data.frame(size=x, size2=x*x)
minSize <- min(Sp1$size, na.rm=TRUE) - 1
maxSize <- max(Sp1$size, na.rm=TRUE) + 2

########
## survival and growth kernels
########
  par(mfrow=c(1,3))
  survModelComp(dataf=Sp1, makePlot=TRUE, legendPos="bottomright")
  so <- makeSurvObj(Sp1, surv ~ size)
  growthModelComp(dataf=Sp1, makePlot=TRUE, legendPos="bottomright")
  go <- makeGrowthObj(Sp1, sizeNext ~ size + size2)
  P.mx <- makeIPMPmatrix(survObj=so, growObj=go, correction="constant",
                            minSize=minSize, maxSize=maxSize)

  image.plot(P.mx@meshpoints, P.mx@meshpoints, t(P.mx), 
             main="P: Survival & growth", xlab="Size (t)", ylab="Size (t+1)")
  abline(a=0, b=1, lty=2)
  diagnosticsPmatrix(P.mx, growObj=go, survObj=so, correction="constant")

########
## fecundity kernel
########
  # flowering
  fo1 <- makeFecObj(Sp1, Formula=fec1Bolt ~ 1, Family="binomial")
  fo2 <- makeFecObj(Sp1, Formula=fec1Bolt ~ size, Family="binomial")
  fo3 <- makeFecObj(Sp1, Formula=fec1Bolt ~ size + size2, Family="binomial")
  par(mfrow=c(1,1))
  fs <- order(Sp1$size)
  fs.fec <- Sp1$fec1Bolt[fs]
  fs.size <- Sp1$size[fs]
  pfz <- tapply(fs.size, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  ps <- tapply(fs.fec, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), ylim=c(0,1), 
       xlab="Size", ylab="Proportion flowering")
  lines(x, predict(fo1@fitFec[[1]], newdata=x0, type="response"), col="red")
  lines(x, predict(fo2@fitFec[[1]], newdata=x0, type="response"), col="green")
  lines(x, predict(fo3@fitFec[[1]], newdata=x0, type="response"), col="blue")
  legend("topleft", col=2:4, lty=1, xjust=1, bg="white",
         legend=sprintf("%s: %s = %.1f", c("1", "size", "size + size2"), 
                        c("AIC"), c(AIC(fo1@fitFec[[1]]), AIC(fo2@fitFec[[1]]),
                                    AIC(fo3@fitFec[[1]]))))
  # stem production
  fo1 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1), 
                    Family=c("binomial", "poisson"), Transform=c("none", "-1"))
  fo2 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~size), 
                    Family=c("binomial", "poisson"), Transform=c("none", "-1"))
  fo3 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~size+size2), 
                    Family=c("binomial", "poisson"), Transform=c("none", "-1"))
  fs <- order(Sp1$size)
  fs.fec <- Sp1$fec2Stem[fs]
  fs.size <- Sp1$size[fs]
  pfz <- tapply(fs.size, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  ps <- tapply(fs.fec, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), 
       xlab="Size", ylab="Number of stems")
  lines(x, predict(fo1@fitFec[[2]], newdata=x0, type="response")+1, col="red")
  lines(x, predict(fo2@fitFec[[2]], newdata=x0, type="response")+1, col="green")
  lines(x, predict(fo3@fitFec[[2]], newdata=x0, type="response")+1, col="blue")
  legend("topleft", col=2:4, lty=1, xjust=1, bg="white",
         legend=sprintf("%s: %s = %.1f", c("1", "size", "size + size2"), 
                        c("AIC"), c(AIC(fo1@fitFec[[2]]), AIC(fo2@fitFec[[2]]),
                                    AIC(fo3@fitFec[[2]]))))

  # inflorescence count
  fo1 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1, 
                                   fec3Head~1), 
                    Family=c("binomial", "poisson", "poisson"), 
                    Transform=c("none", "-1", "none"))
  fo2 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1, 
                                   fec3Head~size), 
                    Family=c("binomial", "poisson", "poisson"), 
                    Transform=c("none", "-1", "none"))
  fo3 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1, 
                                   fec3Head~size+size2), 
                    Family=c("binomial", "poisson", "poisson"), 
                    Transform=c("none", "-1", "none"))
  fs <- order(Sp1$size)
  fs.fec <- Sp1$fec3Head[fs]
  fs.size <- Sp1$size[fs]
  pfz <- tapply(fs.size, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  ps <- tapply(fs.fec, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), 
       xlab="Size", ylab="Number of stems")
  lines(x, predict(fo1@fitFec[[3]], newdata=x0, type="response"), col="red")
  lines(x, predict(fo2@fitFec[[3]], newdata=x0, type="response"), col="green")
  lines(x, predict(fo3@fitFec[[3]], newdata=x0, type="response"), col="blue")
  legend("topleft", col=2:4, lty=1, xjust=1, bg="white",
         legend=sprintf("%s: %s = %.1f", c("1", "size", "size + size2"), 
                        c("AIC"), c(AIC(fo1@fitFec[[3]]), AIC(fo2@fitFec[[3]]),
                                    AIC(fo3@fitFec[[3]]))))

  # final fecundity kernel
  fo <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1, 
                                  fec3Head~size), 
                   Family=c("binomial", "poisson", "poisson"), 
                   Transform=c("none", "-1", "none"),
                   fecConstants=data.frame(seedsPerHead=50,
                                           seedlingEstablishmentRate=0.02))
  F.mx <- makeIPMFmatrix(fecObj=fo, minSize=minSize, maxSize=maxSize)

  image.plot(F.mx@meshpoints, F.mx@meshpoints, t(F.mx), 
             main="F: Sexual reproduction", xlab="Size (t)", ylab="Size (t+1)")

  sens.out <- sensParams(growObj=go, survObj=so, fecObj=fo, 
                         nBigMatrix=50, minSize=minSize, maxSize=maxSize)
  barplot(sens.out$elas, las=2, cex.names=0.5)
```

```{r App_C}
########
## set up
########
  data("dataIPMpackHypericumCov"); help("dataIPMpackHypericumCov")
  d <- dataIPMpackHypericumCov
  d1 <- subset(d, !is.na(d$size) | d$ontogenyNext==1)
  d1 <- subset(d1, d1$year %in% c(1994, 1996))
  fec2 <- 13.78  # nSeeds per fruit
  fec3 <- 0.001336  # pr(germinate_t)
  fec4 <- 0.14  # surv(seedling_t_to_t+1)
  goSB <- 0.08234528
  staySB <- 0.671
  d1 <- d1[,c("year", "size", "surv", "sizeNext", "fec0", "fec1")]

  # assign surviving individuals to 'continuous' stage
  d1$stageNext <- d1$stage <- "continuous"
  d1$stage[is.na(d1$size)] <- NA
  d1$stageNext[d1$surv==0] <- "dead"

  # structure for transition between discrete (SB) & continuous (stem) stages
  d1$number <- 1 
  seedbank <- data.frame(year="All", 
                         size=NA, surv=1, sizeNext=NA,
                         fec0=NA, fec1=NA, 
                         stage=c("seedbank", "seedbank", "continuous"),
                         stageNext=c("seedbank", "continuous", "seedbank"),
                         number=c(staySB, (1-staySB)*fec3*fec4, 1))
  d1 <- rbind(d1, seedbank)
  d1$stage <- as.factor(d1$stage)
  d1$stageNext <- as.factor(d1$stageNext)

########
## construct kernels
########
  # 1994
  d94 <- subset(d1, d1$year %in% c("1994", "All"))
  minSize <- min(d1$size, na.rm=TRUE)
  maxSize <- max(d1$size, na.rm=TRUE)
  x <- seq(minSize, maxSize, length.out=1001)
  x0 <- data.frame(size=x, size2=x^2, size3=x^3)

  # survival
  survModelComp(dataf=d94[!is.na(d$size),], 
                expVars=c(surv ~ 1, 
                          surv ~ size, 
                          surv ~ size + size2, 
                          surv ~ size + size2 + size3), 
                makePlot=TRUE, legendPos="bottomleft", mainTitle="Survival")
  so94 <- makeSurvObj(d94, surv ~ size + I(size^2))

  # growth
  growthModelComp(dataf=d94, 
                  expVars=c(sizeNext ~ 1, 
                            sizeNext ~ size, 
                            sizeNext ~ size + size2, 
                            sizeNext ~ size + size2 + size3),
                  makePlot=TRUE, legendPos="bottomleft", mainTitle="Growth")
  abline(a=0, b=1, lty=2)
  go94 <- makeGrowthObj(d94, sizeNext ~ size)

  # P matrix
  P.mx.94 <- makeIPMPmatrix(survObj=so94, growObj=go94, 
                            minSize=minSize, maxSize=maxSize, 
                            nBigMatrix=80, correction="constant")
  image.plot(P.mx.94@meshpoints, P.mx.94@meshpoints, t(P.mx.94),
             main="P.mx: Survival & growth", xlab="Size (t)", ylab="Size (t+1)")
  abline(a=0, b=1, lty=2, col="white", lwd=2)
  diagnosticsPmatrix(P.mx.94, growObj=go94, survObj=so94, correction="constant")

  # fecundity
  ## flowering probability
  fec0.0_94 <- makeFecObj(d94, Formula=fec0~1, Family="binomial")
  fec0.1_94 <- makeFecObj(d94, Formula=fec0~size, Family="binomial")
  fec0.2_94 <- makeFecObj(d94, Formula=fec0~size+size2, Family="binomial")
  fec0.3_94 <- makeFecObj(d94, Formula=fec0~size+size2+size3, Family="binomial")
  fs <- order(d94$size); fs.fec0 <- d94$fec0[fs]; fs.size <- d94$size[fs]
  pfz <- tapply(fs.size, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  ps <- tapply(fs.fec0, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), xlab="Size", ylab="Pr(flower)")
    y0 <- predict(fec0.0_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0)/(exp(y0) + 1), col=2)
    y0 <- predict(fec0.1_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0)/(exp(y0) + 1), col=3)
    y0 <- predict(fec0.2_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0)/(exp(y0) + 1), col=4)
    y0 <- predict(fec0.3_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0)/(exp(y0) + 1), col=5)
    legend("bottomright", col=2:5, lty=1, xjust=1, bg="white",
           legend=sprintf("%s: %s = %.1f", 
                          c("1","size","size+size2","size+size2+size3"), 
                          c("AIC"), 
                          c(AIC(fec0.0_94@fitFec[[1]]), 
                            AIC(fec0.1_94@fitFec[[1]]), 
                            AIC(fec0.2_94@fitFec[[1]]), 
                            AIC(fec0.3_94@fitFec[[1]]))))
  ## fruit number per plant
  fec1.0_94 <- makeFecObj(d94, Formula=fec1~1, Family="poisson")
  fec1.1_94 <- makeFecObj(d94, Formula=fec1~size, Family="poisson")
  fec1.2_94 <- makeFecObj(d94, Formula=fec1~size+size2, Family="poisson")
  fec1.3_94 <- makeFecObj(d94, Formula=fec1~size+size2+size3, Family="poisson")
  fs.fec1 <- d94$fec1[fs]
  ps <- tapply(fs.fec1, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), xlab="Size", ylab="Fruits per plant")
    y0 <- predict(fec1.0_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0), col=2)
    y0 <- predict(fec1.1_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0), col=3)
    y0 <- predict(fec1.2_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0), col=4)
    y0 <- predict(fec1.3_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0), col=5)
    legend("topleft", col=2:5, lty=1, xjust=1, bg="white",
           legend=sprintf("%s: %s = %.1f", 
                          c("1","size","size+size2","size+size2+size3"), 
                          c("AIC"), 
                          c(AIC(fec1.0_94@fitFec[[1]]), 
                            AIC(fec1.1_94@fitFec[[1]]), 
                            AIC(fec1.2_94@fitFec[[1]]), 
                            AIC(fec1.3_94@fitFec[[1]]))))

  # F matrix
  fo94 <- makeFecObj(d94, 
                     Formula=c(fec0~size+size2+size3, fec1~size+size2+size3),
                     Family=c("binomial", "poisson"), Transform=c("none", -1),
                     meanOffspringSize=mean(d94[is.na(d94$size) & 
                                                  !is.na(d94$sizeNext), 
                                                "sizeNext"]),
                     sdOffspringSize=sd(d94[is.na(d94$size) & 
                                              !is.na(d94$sizeNext), 
                                            "sizeNext"]),
                     fecConstants=data.frame(fec2=fec2, fec3=fec3, fec4=fec4),
                     offspringSplitter=data.frame(seedbank=goSB,
                                                  continuous=(1-goSB)),
                     vitalRatesPerOffspringType=data.frame(
                       seedbank=c(1,1,1,0,0),
                       continuous=c(1,1,1,1,1),
                       row.names=paste0("fec", 0:4)))
  F.mx.94 <- makeIPMFmatrix(fecObj=fo94, minSize=minSize, maxSize=maxSize,
                            nBigMatrix=80, correction="constant")
  image.plot(F.mx.94@meshpoints, F.mx.94@meshpoints, t(F.mx.94[-1, -1]), 
             main="F.mx: Fecundity", xlab="Size (t)", ylab="Size (t-1)")

  # Stage transitions: discrete & continuous
  dto94 <- makeDiscreteTrans(d94)
  dummy94 <- as.matrix(fo94@offspringRel$coefficients[1], list(1, "seedbank"))
  dto94@meanToCont <- as.matrix(dummy94, dimnames=c(1, "seedbank"))
  dummy94 <- as.matrix(fo94@sdOffspringSize, dimnames=list(1, "seedbank"))
  dto94@sdToCont <- as.matrix(dummy94, dimnames=c(1, "seedbank"))
  dto94@discreteTrans[1,1] <- staySB + (1-staySB)*fec3*fec4
  P.mx.94 <- makeIPMPmatrix(growObj=go94, survObj=so94, discreteTrans=dto94,
                            minSize=minSize, maxSize=maxSize, nBigMatrix=80,
                            correction="constant")
  image.plot(P.mx.94@meshpoints, P.mx.94@meshpoints, t(P.mx.94[1:80, 1:80]),
             main="P.mx: Survival & growth", xlab="Size (t)", ylab="Size (t+1)")
  abline(a=0, b=1, lty=2, col="white", lwd=2)

  # Full kernel
  IPM94 <- P.mx.94 + F.mx.94
  Re(eigen(IPM94)$value[1])
  sens.94 <- sens(IPM94)
  image.plot(0:80, 0:80, t(sens.94),
             main="Sensitivity", xlab="Size (t)", ylab="Size (t+1)")
  elas.94 <- elas(IPM94)
  image.plot(0:80, 0:80, t(elas.94),
             main="Elasticity", xlab="SB & Size (t)", ylab="SB & Size (t+1)")
```

```{r App_E.1}
########
## simulate IPM matrices
########
  dff <- generateData()
  gr1 <- makeGrowthObj(dff, sizeNext ~ size + size2)
  sv1 <- makeSurvObj(dff, surv ~ size + size2)
  fv1 <- makeFecObj(dff, Transform="log")
  Pmat <- makeIPMPmatrix(nBigMatrix=100, minSize=-3, maxSize=14,
                         growObj=gr1, survObj=sv1, correction="constant")
  Fmat <- makeIPMFmatrix(nBigMatrix=100, minSize=-3, maxSize=14,
                         fecObj=fv1, correction="constant")
  IPM <- Pmat
  IPM@.Data <- Pmat+Fmat  # retains Pmat meshpoints, etc

  w.eigen <- Re(eigen(IPM)$vectors[,1])
  stable.dist <- w.eigen/sum(w.eigen)

########
## generate starting population
#########
  # size distribution
  size0 <- runif(1000, min(IPM@meshpoints), max(IPM@meshpoints))
  # IPM breakpoints = (mdpts - h/2):(max(mdpts) + h/2)
  breakpoints <- c(IPM@meshpoints - (IPM@meshpoints[2] - IPM@meshpoints[1])/2,
                   IPM@meshpoints[length(IPM@meshpoints)] +
                     (IPM@meshpoints[2] - IPM@meshpoints[1])/2)
  # bin size0 based on breakpoints
  new.dist.i <- findInterval(size0, breakpoints, all.inside=TRUE)
  loc.sizes <- table(new.dist.i)
  new.dist.n0 <- rep(0, length(IPM@meshpoints))
  new.dist.n0[as.numeric(names(loc.sizes))] <- loc.sizes

########
## project population forward
########
  tmax <- 10
  storage <- matrix(NA, nrow=dim(Pmat)[1], ncol=tmax)
  new.dist.n <- new.dist.n0
  for(k in 1:tmax) {
    new.dist.n <- IPM %*% new.dist.n
    storage[,k] <- new.dist.n
  }

########
## plots
########
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  image.plot(Pmat@meshpoints, Pmat@meshpoints, t(Pmat), 
             main="P", xlab="Size (t)", ylab="Size (t+1)")
  contour(Pmat@meshpoints, Pmat@meshpoints, t(Pmat), add=TRUE)
  image.plot(Pmat@meshpoints, Pmat@meshpoints, t(Fmat), 
             main="F", xlab="Size (t)", ylab="")
  contour(Pmat@meshpoints, Pmat@meshpoints, t(Fmat), add=TRUE)
  hist(size0, col="grey", xlim=range(Pmat@meshpoints), freq=FALSE,
       xlab="Size (t=0)", ylab="", main="")
  for(k in 1:tmax) {
    lines(Pmat@meshpoints, storage[,k]/sum(storage[,k]), col=rgb(0,0,k/tmax,1))
  }
  lines(Pmat@meshpoints, stable.dist, col="red")
  image.plot(1:tmax, Pmat@meshpoints, t(log(storage)), 
             xlab="time", ylab="Size")
  abline(v=1.5+(1*c(0:tmax)))
  Re(eigen(IPM)$values[1])
```

```{r App_E.2}
dff <- generateData()

# good year
gr1 <- makeGrowthObj(dff, Formula=sizeNext ~ size, 
                     regType="constantVar", Family="gaussian")
sv1 <- makeSurvObj(dff, Formula=surv ~ size + size2)
fv1 <- makeFecObj(dff, Formula=list(fec~size), Transform="log")
fv1@fitFec[[1]]$coefficients <- c(-1, 0.35)

# bad year
sv1 <- coerceSurvObj(survObj=sv1, coeff=c(-1, 1, 2))
sv2 <- coerceSurvObj(survObj=sv1, coeff=c(-3, 0, -2))
gr2 <- coerceGrowthObj(gr1, coeff=c(-3, -3), sd=1)

# kernel
Pmat1 <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
                        growObj=gr1, survObj=sv1)
Pmat2 <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
                        growObj=gr2, survObj=sv2)
Fmat <- makeIPMFmatrix(nBigMatrix=50, minSize=-2, maxSize=12, fecObj=fv1)
IPM1 <- Pmat1 + Fmat
IPM2 <- Pmat2 + Fmat

# plots
image.plot(Pmat1@meshpoints, Pmat1@meshpoints, t(Pmat1), main="good year")
image.plot(Pmat1@meshpoints, Pmat1@meshpoints, t(Pmat2), main="bad year")

exp(stochGrowthRateSampleList(list(IPM1, IPM2), nRunIn=1000, tMax=5000))

nsim <- 100
tmax <- 100
IPM.ls <- list(IPM1, IPM2)
nt0 <- abs(Re(eigen(IPM1)$vector[,1])*10)
rt <- rtGood <- rtBad <- matrix(NA, nsim, tmax)
for(j in 1:nsim) {
  ntGood <- ntBad <- nt <- nt0
  for(k in 1:tmax) {
    sv.t <- coerceSurvObj(sv1, coeff=runif(3, -3, 3))
    gr.t <- coerceGrowthObj(gr1, coeff=runif(2, -3, 3), sd=1)
    Pmat.t <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
                        growObj=gr.t, survObj=sv.t)
    IPM.t <- Pmat.t + Fmat
    nt <- IPM.t %*% nt
    rt[j,k] <- sum(nt)
    ntGood <- IPM1 %*% ntGood
    rtGood[j,k] <- sum(ntGood)
    ntBad <- IPM2 %*% ntBad
    rtBad[j,k] <- sum(ntBad)
  }
}

par(mfrow=c(1,1), bty="l")
matplot(t(rt), type="l", xlab="time", ylab="N", log="y", 
        ylim=c(range(c(rtGood, rtBad))), lty=1, col=rgb(0,0,0,0.5))
points(rtGood[1,], type="l", lwd=2, lty=2)
points(rtBad[1,], type="l", lwd=2, lty=2)
```

```{r App_E.3}
dff <- generateData()
go <- makeGrowthObj(dff, sizeNext ~ size + size2 + covariate)
so <- makeSurvObj(dff, surv ~ size + size2)
eo <- makeEnvObj(dff)
fo <- makeFecObj(dff, Formula=list(fec~size))

Pmat <- makeCompoundPmatrix(nEnvClass=2, envMatrix=eo, 
                            minSize=-3, maxSize=14, nBigMatrix=100,
                            growObj=go, survObj=so, 
                            correction="constant")
image.plot(1:nrow(Pmat), 1:ncol(Pmat), t(log(Pmat)), axes=F, 
           xlab="Size (t)", ylab="Size (t+1)")
axis(1, at=1:nrow(Pmat), lab=round(rep(Pmat@meshpoints, Pmat@nEnvClass)))
axis(2, at=1:nrow(Pmat), lab=round(rep(Pmat@meshpoints, Pmat@nEnvClass)))
abline(h=length(Pmat@meshpoints)*(1:Pmat@nEnvClass))
abline(v=length(Pmat@meshpoints)*(1:Pmat@nEnvClass))
```

```{r App_F}
########
## load data
########
  load("data/ActaeaRegDat.RData")
  d <- ActaeaRegDat

########
## regressions
########
  g.lm <- lm(y ~ x + K, data=d)
  s.lm <- glm(surv ~ x, family="binomial", data=d)
  fl.lm <- glm(flow ~ x, family="binomial", data=d)
  fr.lm <- glm(fruits ~ x, family="poisson", data=d)

  par(mfrow=c(2,2), mar=c(4,4,2,1))
  xx <- data.frame(x=seq(2, 14, length.out=100), K=25)
  plot(d$x, d$surv, col=rgb(0,0,0,0.1), xlab="Size (t)", ylab="Survival to t+1")
  lines(xx$x, predict(s.lm, xx, type="response"), col="red", lwd=2)
  plot(d$x, d$y, xlab="Size (t)", ylab="Size (t+1)")
  lines(xx$x, predict(g.lm, xx, type="response"), col="red", lwd=2)
  plot(d$x, d$flow, col=rgb(0,0,0,0.1), xlab="Size (t)", ylab="Pr(Flower)")
  lines(xx$x, predict(fl.lm, xx, type="response"), col="red", lwd=2)
  plot(d$x, d$fruits, xlab="Size (t)", ylab="Fruit number")
  lines(xx$x, predict(fr.lm, xx, type="response"), col="red", lwd=2)

  p <- rep(0, 15)
  p <- list("s.b0"=coef(s.lm)[1], "s.bx"=coef(s.lm)[2], 
         "g.b0"=coef(g.lm)[1], "g.bx"=coef(g.lm)[2], "g.bK"=coef(g.lm)[3],
         "fl.b0"=coef(fl.lm)[1], "fl.bx"=coef(fl.lm)[2],
         "fr.b0"=coef(fr.lm)[1], "fr.bx"=coef(fr.lm)[2],
         "est.x.mn"=3.0721075, "est.x.sd"=1.448899,
         "sd.xy"=1.155625, 
         "est.sdlg"=0.0091550, "surv.sdlg"=0.2383389, 
         "seed.pred"=0.195905)

########
## define functions
########
  # survival probability
  sx <- function(x, p) {
    u <- exp(p$s.b0 + p$s.bx*x)
    return(u/(1+u))
  }

  # growth
  gyx <- function(y, x, K, p) {
    dnorm(y, p$g.b0 + p$g.bx*x + p$g.bK*K, p$sd.xy)
  }

  # fecundity
  fx <- function(x, p) {
    # flowering probability
    u <- exp(p$fl.b0 + p$fl.bx*x)
    # fruits per flowering individual
    n.fr <- exp(p$fr.b0 + p$fr.bx*x)
    n.fr[n.fr<0] <- 0
    n.seed.per.fruit <- p$seed.pred*0.38901 + (1 - p$seed.pred)*9.61368
    return(u/(1+u)*n.fr*n.seed.per.fruit)
  }

  # seedling survival and size distribution
  py3 <- function(y, p) {
    return(p$surv.sdlg * dnorm(y, p$est.x.mn, p$est.x.sd)/
             (1-pnorm(minsize, p$est.x.mn, p$est.x.sd)))
  }

  # iterator function
  iter <- function(Nt) {
    Nt1 <- array(0, dim=c(2+n.size)) 
    Nt1[1] <- Nt1[1] + sum(FF %*% Nt[3:(n.size+2)])
    Nt1 <- Nt1 + as.vector(P %*% Nt)
    return(Nt1)
  }

########
## construct kernel
########
  # define breakpoints, meshpoints, stepsize
  n.size <- 100
  tol <- 1.e-8
  minsize <- 0.9234274; maxsize <- 15.27885
  L <- minsize; U <- maxsize; n <- n.size
  b <- L + c(0:n)*(U-L)/n
  y <- 0.5*(b[1:n]+b[2:(n+1)])
  h <- y[2]-y[1]

  # loop over forest succession (=approx(cells))
  spruce <- 0
  K.conc <- 0
  for(k in 1:120) {
    # spruce proportion & K concentration
    spruce <- 1/(1+exp(-0.09*k+5))
    K.conc <- exp(3.00932 - 0.53267*spruce)*1.722324
    # kernel components
      ## growth (including eviction correction)
      G <- array(0, dim=c(n.size, n.size))
      G <- h*outer(y, y, gyx, p=p, K=K.conc)
      for(i in 1:(n.size/2)) G[1,i] <- G[1,i] + 1 - sum(G[,i])
      for(i in (n.size/2+1):n.size) G[n.size,i] <- G[n.size,i] + 1 - sum(G[,i])
      ## survival
      S <- sx(y, p)
      FF <- fx(y, p)*h
      P <- array(0, dim=c(2+n.size, 2+n.size)) # [1,] = seed; [2,] = seedling
      P[2,1] <- p$est.sdlg
      P[3:(2+n.size),2] <- py3(y, p)
      for(i in 3:(2+n.size)) P[i,3:(2+n.size)] <- S*G[(i-2),]

    # population iterations
    Nt <- array(1, dim=c(2+n.size))
    Nt <- Nt/sum(Nt)
    qmax <- 1000; lam <- 1
    while(qmax > tol) {
      Nt1 <- iter(Nt)
      qmax <- sum(abs(Nt1 - lam*Nt))
      lam <- sum(Nt1)
      Nt <- Nt1/lam
    }
    stable.dist <- Nt/sum(Nt)
    stable.lam <- lam
    cat("Year:", k, "Proportion spruce:", round(spruce,2), " Lambda:", lam, "\n")
  }
```