### 2018-03-13 exploration.Rmd
added more covariate plots

---
title: "Exploration"
author: "Tim Szewczyk"
date: "2/16/2018"
output: html_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = T)
pkgs <- c("fields", "sevcheck", "gbPopMod", "IPMpack", "tidyverse", "magrittr")
suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
theme_set(theme_bw())
source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
lc.df <- read_csv("data/landcover_5km.csv") %>% 
  filter(!is.na(bio1_mean)) %>%
  mutate(x=as.integer(factor(.$left)),
         y=as.integer(factor(.$top,
                             levels=rev(levels(factor(.$top))))),
         x_y=paste(x, y, sep="_"))
```

# Exploration of options for the simulation component

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
mean_abund <- 200
regPlots <- T
ipmPlots <- T
nPlots <- 5
z.rng <- c(1, 12)
n <- 50  # ncells in IPM matrix
tmax <- 50
cells <- 1:n_distinct(lc.df$CellID)
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-8, 2.1, -.09), # b1 + b2*z + b3*z^2
                    s_x=c(2, -.1, -2, -.1, 2, -2, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(2, -.1, 2, -.1, 2, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(-2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, -1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.03, # p(establishment)
                    NDD=T, # negative density dependence: p_est=p_est/N
                    NDD_n=mean_abund/2, # mean number of recruits if NDD
                    rcr_SB=0.5, # p(recruit from seedbank)
                    rcr_dir=0.5, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    sdd_max=5, # max SDD distance in cells
                    sdd_rate=1 # SDD dispersal rate
                    ))
sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
```

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
                                 y=1:max(lc.df$y))) %>%
  mutate(x_y=paste(x, y, sep="_")) %>%
  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
         temp2=temp^2,
         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
         prec2=prec^2,
         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
         inbd=!is.na(match(.$x_y, lc.df$x_y)),
         id=row_number(), 
         id.inbd=min_rank(na_if(inbd*id, 0)),
         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
env.rct[is.na(env.rct)] <- 0
sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
                        g.p=list(sdd.max=sp_i$p$sdd_max, 
                                 sdd.rate=sp_i$p$sdd_rate, 
                                 bird.hab=rep(1,5)))
env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
rm(env.rct)
# sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
# ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,740])) + geom_tile()
```

---

## Simulation: IPM framework
```{r S}
########
## set up
########
# set up storage objects & environmental covariates
p <- sp_i$p
sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
sdd[,,1:2,] <- sdd.pr
n_z <- list(s=length(p$s_z), g=length(p$g_z),
            fl=length(p$fl_z), seed=length(p$seed_z))
n_x <- list(s=length(p$s_x), g=length(p$g_x),
            fl=length(p$fl_x), seed=length(p$seed_x))
X <- list(s=as.matrix(env.df[,1:n_x$s]), g=as.matrix(env.df[,1:n_x$g]),
          fl=as.matrix(env.df[,1:n_x$fl]), seed=as.matrix(env.df[,1:n_x$seed]))

########
## generate underlying IPM and simulated data
########

#---- Truth: use assigned slopes to fill IPM matrix
U <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.75, discrete=1,
                       p, n_z, n_x, X, sdd.pr, env.df, verbose=T)

#---- Realization: generate simulated data
S <- simulate_data(n.cell, tmax, n_i, z.rng, U$lo, U$hi, 
                   p, X, n_z, sdd, U$sdd.j, verbose=T)

#---- Aggregate results
lam.df <- env.df
lam.df$lambda <- apply(U$IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% 
  add_column(lam.S=S$N_sim[,tmax]/(S$N_sim[,tmax-1]+.01),
             nSeed=S$nSd[,tmax], D=S$D[,tmax], 
             B0=S$B[,1], Btmax=S$B[,tmax+1], 
             N.S=map_dbl(S$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
             Surv.S=map_dbl(S$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
             Rcr.S=map_dbl(S$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
  mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
  add_column(N.U=apply(U$Nt[-1,,tmax+1],2,sum), lam.U=U$lam.t[,tmax])

########
## plots
########
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lambda))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.U))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.S>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>0)) + geom_tile() 
ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>0 & lam.U<1)) + geom_tile() 
ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0 & lambda<1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.U))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(Surv.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(Rcr.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=nSeed)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax>B0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
par(mfrow=c(3,3))
plot(log(lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pOth, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pDec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pEvg, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pMxd, data=lam.df); abline(h=0, col="red", lty=2)
plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
plot(density(log(lam.df$lam.U))); abline(v=0, col="red", lty=2)
# matplot(t(log(U$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
# matplot(t(log(apply(U$Nt[-1,,], 2:3, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
# matplot(t(log(S$N_sim[,2:tmax]/(S$N_sim[,1:(tmax-1)]))), 
#         type="l", lty=1, col=rgb(0,0,0,.1))
# matplot(t(log(S$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
#matplot(t(log(nSd*(1-p$p_emig)+D)), type="l", lty=1, col=rgb(0,0,0,0.1))

sample.cells <- sample((1:n.cell)[lam.df$Surv.S>0], nPlots)
if(regPlots) {
  par(mfrow=c(3,2), mar=c(4,4,2,1))
  xx <- tibble(size=seq(U$lo, U$hi, length.out=100), size2=size^2, size3=size^3)
  E.t <- xx %>% add_column(s=NA, g=NA, fl=NA, seed=NA)
  xxx <- seq(0, U$hi*.5, length.out=100)
  for(i in sample.cells) {
    z.t <- S$d[[i]][!is.na(S$d[[i]]$size) & S$d[[i]]$yr==tmax,]
    E.t <- sim_expected(E.t, 1:100, 1:100, E.t, p, map(X, ~.[i,]), n_z)
    main.i <- paste("Cell", i)
    w1 <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      lines(E.t$size, E.t$s, col=2)
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(U$lo,U$hi),
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      lines(E.t$size, E.t$g, col=2)
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      lines(E.t$size, E.t$fl, col=2)
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      lines(E.t$size, E.t$seed, col=2)
    plot(U$y, stable.dist1, ylab="", type="l", col=2, 
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    hist(with(S$d[[i]], sizeNext[is.na(size)]), col="grey", freq=F, 
         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red")
  }
}
if(ipmPlots) {
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  x <- c(0, U$y)
  for(i in sample.cells) {
   z.t <- S$d[[i]][!is.na(S$d[[i]]$size) & S$d[[i]]$yr==tmax,]
   main.i <- paste("Cell", i)
   z.t$y <- U$y[findInterval(z.t$size, U$y, all.inside=T)]
   totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
   image.plot(x, x, t(U$Ps[,,i]), main=paste("P matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
    abline(a=0, b=1, col="white", lty=2)
   image.plot(x, x, t(U$Fs[,,i]), main=paste("F matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*U$hi*.9, col="white")
   image.plot(x, x, t(sens(U$IPMs[,,i])), main=paste("Sensitivity:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
   image.plot(x, x, t(elas(U$IPMs[,,i])), main=paste("Elasticity:", main.i),
              xlab="Size (t)", ylab="Size (t+1)")
   points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
  }
}
par(mfrow=c(2,2), mar=c(4,4,2,1))
x <- c(0, U$y)
image.plot(x, x, t(apply(U$Ps, 1:2, mean)), main="P matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, t(apply(U$Fs, 1:2, mean)), main="F matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, sens(t(apply(U$IPMs, 1:2, mean))), main="Sensitivity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, elas(t(apply(U$IPMs, 1:2, mean))), main="Elasticity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
```

----

## Observed datasets
The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data. The realized simulated data generated above (`d` and its abundance summary `N_sim`) are the underlying true distribution. The samples will be drawn from occupied cells in the final time steps where the species is approximately at equilibrium.
```{r O_setup}
O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
```

### Correlative: MaxEnt
```{r O_Mx}
library(dismo)
Mx.P <- sample(P.i, O_n$Mx, replace=T)
pres <- 1:nrow(lam.df) %in% Mx.P
Mx.out <- maxent(x=lam.df[,c(1,3,5:9)], p=pres)
lam.df$MxPred <- 1:nrow(lam.df) %in% row.names(Mx.out@presence)
lam.df$MxCorrect <- with(lam.df, (MxPred==1 & lambda > 1) |
                           (MxPred==0 & lambda < 1))
ggplot(lam.df, aes(x=lon, y=lat, fill=MxCorrect)) + geom_tile()
```
### Mechanistic: CA
```{r O_CA}

```
### Mechanistic: IPM
```{r O_IPM}
IPM.P <- sample(P.i, O_n$IPM, replace=FALSE)
IPM.d <- vector("list", O_n$IPM)
for(i in IPM.P) {
  IPM.d[[i]] <- S$d[[i]] %>% filter(yr %in% O_yr$IPM) %>%
    add_column(id.inbd=i) %>%
    full_join(env.df[i,], by="id.inbd")
}
IPM.d <- do.call(rbind, IPM.d)
p.ipm <- p
```

---

## Fitting the models
Each observed dataset *O* is used to predict the distribution of species *S*. The predicted distribution *P* is compared to the known truth *S* to assess the predictive accuracy of each modeling approach, given any limitations, uncertainty, or bias imposed in *O*. 
### Correlative: MaxEnt
```{r P_Mx}

```
### Mechanistic: CA
```{r P_CA}

```
### Mechanistic: IPM
```{r P_IPM}
# run chunk O_IPM first to generate observed dataset `IPM.d`

########
## fit models
########
s.m <- glm(surv ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn + 
               pOth + pDec, data=IPM.d, family="binomial")
g.m <- lm(sizeNext ~ size + I(size^2) + temp + temp2 + prec + prec2 +
              pOpn + pOth, data=IPM.d)
fl.m <- glm(fl ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn +
                pOth, data=IPM.d, family="binomial")
seed.m <- glm(seed ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn, 
              data=IPM.d, family="poisson")
p.ipm$s_z <- coef(s.m)[1:length(p.ipm$s_z)]
p.ipm$s_x <- coef(s.m)[(1:length(p.ipm$s_x)) + length(p.ipm$s_z)]
p.ipm$g_z <- coef(g.m)[1:length(p.ipm$g_z)]
p.ipm$g_x <- coef(g.m)[(1:length(p.ipm$g_x)) + length(p.ipm$g_z)]
p.ipm$g_sig <- summary(g.m)$sigma
p.ipm$fl_z <- coef(fl.m)[1:length(p.ipm$fl_z)]
p.ipm$fl_x <- coef(fl.m)[(1:length(p.ipm$fl_x)) + length(p.ipm$fl_z)]
p.ipm$seed_z <- coef(seed.m)[1:length(p.ipm$seed_z)]
p.ipm$seed_x <- coef(seed.m)[(1:length(p.ipm$seed_x)) + length(p.ipm$seed_z)]
p.ipm$rcr_z <- c(mean(IPM.d[is.na(IPM.d$size), "sizeNext"]),
                 sd(IPM.d[is.na(IPM.d$size), "sizeNext"]))

n_z <- list(s=length(p.ipm$s_z), g=length(p.ipm$g_z),
            fl=length(p.ipm$fl_z), seed=length(p.ipm$seed_z))
n_x <- list(s=length(p.ipm$s_x), g=length(p.ipm$g_x),
            fl=length(p.ipm$fl_x), seed=length(p.ipm$seed_x))
X <- list(s=as.matrix(env.df[,1:n_x$s]), g=as.matrix(env.df[,1:n_x$g]),
          fl=as.matrix(env.df[,1:n_x$fl]), seed=as.matrix(env.df[,1:n_x$seed]))

########
## predict distribution with fit models
########
#---- Truth: use estimated slopes to fill IPM matrix
U.f <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
                         p.ipm, n_z, n_x, X, sdd, env.df, verbose=T)

#---- Realization: use estimated slopes to generate simulated data
S.f <- simulate_data(n.cell, tmax, n_i, z.rng, U.f$lo, U.f$hi, 
                     p.ipm, X, n_z, sdd, U.f$sdd.j, verbose=T)

#---- Aggregate results
lam.df$lambda.f <- apply(U.f$IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% 
  add_column(lam.S.f=S.f$N_sim[,tmax]/(S.f$N_sim[,tmax-1]+.01),
             nSeed.f=S.f$nSd[,tmax], D.f=S.f$D[,tmax], 
             B0.f=S.f$B[,1], Btmax.f=S.f$B[,tmax+1], 
             N.S.f=map_dbl(S.f$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
             Surv.S.f=map_dbl(S.f$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
             Rcr.S.f=map_dbl(S.f$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
  mutate(nSdStay.f=nSeed.f*(1-p.ipm$p_emig), nSdLeave.f=nSeed.f*p.ipm$p_emig) %>%
  add_column(N.U.f=apply(U.f$Nt[,,tmax],2,sum), lam.U.f=U.f$lam.t[,tmax-1])

########
## plots
########
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=lambda.f-lambda, colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black")) +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=lam.S.f-lam.S, colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black")) +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=lam.U.f-lam.U, colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black")) +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=lam.S>1, colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black"))
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=lam.S.f>1, colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black"))
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=(lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1),
                colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black"))
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=(lam.S>1 & lam.S.f>1) | (lam.S<1 & lam.S.f<1),
                colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black"))
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=(N.S>1 & N.S.f>1) | (N.S<1 & N.S.f<1),
                colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black"))
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=(lam.U>1 & lam.U.f>1) | (lam.U<1 & lam.U.f<1),
                colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black"))
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=(N.U>1 & N.U.f>1) | (N.U<1 & N.U.f<1),
                colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black"))
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=(N.S.f-N.S)/N.S, colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black")) +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes( fill=(Surv.S.f-Surv.S)/Surv.S, colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black")) +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat)) + 
  geom_tile(aes(fill=(Rcr.S.f-Rcr.S)/Rcr.S, colour=id.inbd %in% IPM.P)) +
  scale_colour_manual(values=c(NA, "black")) +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
par(mfrow=c(3,3))
plot((lambda.f-lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot((lambda.f-lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot((lambda.f-lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
plot((lambda.f-lambda) ~ pOth, data=lam.df); abline(h=0, col="red", lty=2)
plot((lambda.f-lambda) ~ pDec, data=lam.df); abline(h=0, col="red", lty=2)
plot((lambda.f-lambda) ~ pEvg, data=lam.df); abline(h=0, col="red", lty=2)
plot((lambda.f-lambda) ~ pMxd, data=lam.df); abline(h=0, col="red", lty=2)
plot(density(lambda.f-lambda)); abline(v=0, col="red", lty=2)
plot(density(lam.U.f-lam.U)); abline(v=0, col="red", lty=2)
# matplot(t(log(U.f$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
# matplot(t(log(apply(U.f$Nt, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
# matplot(t(S.f$N_sim[,2:tmax]/S$N_sim[,1:(tmax-1)]), 
#         type="l", lty=1, col=rgb(0,0,0,.1))
# matplot(t(log(S.f$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))

sample.cells <- sample((1:n.cell)[lam.df$Surv.S>0], nPlots)
if(regPlots) {
  par(mfrow=c(3,2), mar=c(4,4,2,1))
  xx <- tibble(size=seq(U.f$lo, U.f$hi, length.out=100), 
               size2=size^2, size3=size^3)
  E.t <- xx %>% add_column(s=NA, g=NA, fl=NA, seed=NA)
  xxx <- seq(0, U.f$hi*.5, length.out=100)
  for(i in sample.cells) {
    z.t <- S.f$d[[i]][!is.na(S.f$d[[i]]$size) & S.f$d[[i]]$yr==tmax,]
    E.t <- sim_expected(E.t, 1:100, 1:100, E.t, p.ipm, map(X, ~.[i,]), n_z)
    E.S.t <- sim_expected(E.t, 1:100, 1:100, E.t, p, map(X, ~.[i,]), n_z)
    main.i <- paste("Cell", i)
    w1 <- Re(eigen(U.f$IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    w.S <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist.S <- w.S/sum(w.S)
    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      lines(E.t$size, E.t$s, col=2); lines(E.S.t$size, E.S.t$s, col=4)
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(U$lo,U$hi),
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      lines(E.t$size, E.t$g, col=2); lines(E.S.t$size, E.S.t$g, col=4)
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      lines(E.t$size, E.t$fl, col=2); lines(E.S.t$size, E.S.t$fl, col=4)
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      lines(E.t$size, E.t$seed, col=2); lines(E.S.t$size, E.S.t$seed, col=4)
    plot(U.f$y, stable.dist1, ylab="", type="l", col=2, 
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
      lines(U.f$y, stable.dist.S, col=4)
    hist(with(S.f$d[[i]], sizeNext[is.na(size)]), col="grey", freq=F, 
         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
      lines(xxx, dnorm(xxx, p.ipm$rcr_z[1], p.ipm$rcr_z[2]), col=2)
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col=4)
  }
}
if(ipmPlots) {
  for(i in sample.cells) {
   z.t <- S.f$d[[i]][!is.na(S.f$d[[i]]$size) & S.f$d[[i]]$yr==tmax,]
   main.i <- paste("Cell", i)
   z.t$y <- U.f$y[findInterval(z.t$size, U.f$y, all.inside=T)]
   totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
   image.plot(x, x, t(U.f$Ps[,,i]), main=paste("P matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
    abline(a=0, b=1, col="white", lty=2)
   image.plot(x, x, t(U.f$Fs[,,i]), main=paste("F matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*U.f$hi*.9, col="white")
   image.plot(x, x, t(sens(U.f$IPMs[,,i])), main=paste("Sensitivity:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
   image.plot(x, x, t(elas(U.f$IPMs[,,i])), main=paste("Elasticity:", main.i),
              xlab="Size (t)", ylab="Size (t+1)")
   points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
  }
}
par(mfrow=c(2,2), mar=c(4,4,2,1))
x <- c(0, U.f$y)
image.plot(x, x, t(apply(U.f$Ps, 1:2, mean)), main="P matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, t(apply(U.f$Fs, 1:2, mean)), main="F matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, sens(t(apply(U.f$IPMs, 1:2, mean))), main="Sensitivity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, elas(t(apply(U.f$IPMs, 1:2, mean))), main="Elasticity: mean",
           xlab="Size (t)", ylab="Size (t+1)")

IPM.fit.sum <- lam.df %>% 
  summarise(mnLamDiff=mean(lambda.f-lambda),
            mdLamDiff=median(lambda.f-lambda),
            sdLamDiff=sd(lambda.f-lambda),
            pctLamCorr=mean((lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1)),
            mnLamSDiff=mean(lam.S.f-lam.S, na.rm=T),
            mdLamSDiff=median(lam.S.f-lam.S, na.rm=T),
            sdLamSDiff=sd(lam.S.f-lam.S, na.rm=T),
            pctLamSCorr=mean((lam.S>1 & lam.S.f>1) | (lam.S<1 & lam.S.f<1)),
            mnNDiff=mean(N.S.f-N.S),
            mdNDiff=median(N.S.f-N.S),
            sdNDiff=sd(N.S.f-N.S),
            pctNCorr=mean((N.S>0 & N.S.f>0) | (N.S==0 & N.S.f==0)),
            mnSurvDiff=mean(Surv.S.f-Surv.S),
            mdSurvDiff=median(Surv.S.f-Surv.S),
            sdSurvDiff=sd(Surv.S.f-Surv.S),
            pctSurvCorr=mean((Surv.S>0 & Surv.S.f>0) | (Surv.S==0 & Surv.S.f==0)),
            mnRcrDiff=mean(Rcr.S.f-Rcr.S),
            mdRcrDiff=median(Rcr.S.f-Rcr.S),
            sdRcrDiff=sd(Rcr.S.f-Rcr.S),
            pctRcrCorr=mean((Rcr.S>0 & Rcr.S.f>0) | (Rcr.S==0 & Rcr.S.f==0)))
str(IPM.fit.sum)
```